---
title: "Data Manipulation in R HW"
author: "Dean D'sozua"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```  
  
  
#### Install the package HSAUR2.  
  + After installing the packages, load the package using the library command.
  + After the library command, pull up the data set CHFLS using the data command.
  + If you're unsure how to do this, try help().
  
```{r, warning=FALSE, message=FALSE}
## Installing package:
# install.packages("HSAUR2")
 
library(HSAUR2)
data(CHFLS)

head(CHFLS)
``` 
#### What are the names of the columns of the CHFLS data set?  
  
  The names of the colums are obtained as follows:  
  
```{r, warning=FALSE, message=FALSE}
names(CHFLS)
```  

#### What is the class of CHFLS?  
  
  The class of CHFLS is obtained as follows:  
  
```{r, warning=FALSE, message=FALSE}
class(CHFLS)
```  

#### What is the average R_income for each R_edu ?
  
  For each level of R_edu we obtain the average as follows:  
  
```{r, warning=FALSE, message=FALSE}
moredu<-tapply(CHFLS$R_income, CHFLS$R_edu, mean)
moredu
``` 
  
#### Create a table of counts of rows by the values in the R_health column and the R_happy column. In other words, how many rows are in the data frame for each combination of distinct values in these two columns?  
  
```{r, warning=FALSE, message=FALSE}
table(CHFLS$R_health,CHFLS$R_happy)
```

#### What is your working directory?  
  
  The working directory can be obtained as follows:  
  
```{r, warning=FALSE, message=FALSE}
getwd()
```  

#### What files exist in your working directory?  
  
  The files contained in the working directory are as follows:  
  
```{r, warning=FALSE, message=FALSE}
list.files()
```   

#### Save the data frame CHFLS as a csv file in your working directory.  Show that CHFLS exists as a csv file  
  
```{r, warning=FALSE, message=FALSE}
# Write the CHFLS data frame as a .csv file in the working directory
write.csv(CHFLS, file = "MyChfls.csv", fileEncoding = "ISO-8859-1")

# Listing all files to verify that it has been written
list.files()
```  
  
  As can be seen from the above list of files, __MyChfls.csv__ file exists in the working directory showing that it was written successfully.  
  We can also read from this file again as follows:

```{r, warning=FALSE,message=FALSE}
temp<-read.csv("MyChfls.csv", header = TRUE, sep = ",", quote = "\"", dec = ".", fill = TRUE, comment.char = "")
head(temp)
```