---
title: "Survival Analysis"
author: "Dean D'souza"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solution:  
  
  The following code chunk displays the setup procedure:  
  
```{r,warning=FALSE,message=FALSE}
#Installing the 'coin' & 'TH.data' package (commented out as not necessary)
#install.packages("coin")
#install.packages("TH.data")

# Loading the data sets
library("coin")
library("TH.data")

data("glioma")
data("GBSG2")

head(glioma)
head(GBSG2)

summary(glioma)
summary(GBSG2)
```
  
* After loading the two data sets, execute the commands in this lecture three times:  
  + First, using glioma and the two groupings based on the sex column  
  
```{r,warning=FALSE,message=FALSE}
# Checking the levels of 'sex' covariate
levels(glioma$sex)

# Comparing time across sexes
table(subset(glioma, sex=="Female", select = time))
table(subset(glioma, sex=="Male", select = time))

# Building the 'Surv' Object
surv1<-Surv(time=glioma$time, event=glioma$event)
summary(surv1)

# Estimating the Survival Curve
survf1<-survfit(formula = surv1~glioma$sex, data=glioma)
summary(survf1)

# Plotting the results
plot(survf1, lty=1:2, main="Glioma (Splitting based on sex)", ylab = "Probability", xlab="Survival Time")
legend("topright",legend=levels(glioma$sex), lty=1:2)
```  
  
  The above calculations and plot shows us that as time increases, initially Males have a higher chance of survival, but later, we observe that Females have a higher chance of survival. This eventually balances out to less than 40% chance of survival after more than 50 months.  
  
  + Second, using glioma and the two groupings based on the histology column  
    
```{r,warning=FALSE,message=FALSE}
# Checking the levels of 'histology' covariate
levels(glioma$histology)

# Comparing time across hoistology levels
table(subset(glioma, histology=="GBM", select = time))
table(subset(glioma, histology=="Grade3", select = time))

# Building the 'Surv' Object
surv2<-Surv(time=glioma$time, event=glioma$event)
summary(surv2)

# Estimating the Survival Curve
survf2<-survfit(formula = surv2~glioma$sex, data=glioma)
summary(survf2)

# Plotting the results
plot(survf2, lty=1:2, main="Glioma (Splitting based on Histology)", ylab = "Probability", xlab="Survival Time")
legend("topright",legend=levels(glioma$histology), lty=1:2)
```  
  
  As seen with the case of splitting based on sex, splitting based on histology leads to similar results with the `GBM` and `Grade3` histologies as seen by the plot above.  
  
  + Third, using the GBSG2 data frame from the TH.data package, using the column horTH to split the data frame into two groups  
  
```{r,warning=FALSE,message=FALSE}
# Checking the levels of 'horTH' covariate
levels(GBSG2$horTh)

# Comparing time across horTH
table(subset(GBSG2, GBSG2$horTH=="no", select = time))
table(subset(GBSG2, GBSG2$horTH=="yes", select = time))

# Building the 'Surv' Object
surv3<-Surv(time=GBSG2$time, event=GBSG2$cens)
summary(surv3)

# Estimating the Survival Curve
survf3<-survfit(formula = surv3~GBSG2$horTh, data=GBSG2)
summary(survf3)

# Plotting the results
plot(survf3, lty=1:2, main="GBSG2 (Split based on horTH)", ylab = "Probability", xlab="Survival Time")
legend("topright",legend=levels(GBSG2$horTh), lty=1:2)
```  
  
  From the above plot and calculations we can see that as time progresses, those who receive hormonal therapy have a better chance at survival, with those not recieving the therapy having a similarly steady decrease with a sudden drop in survival after roughly 2450 days.  
  

