---
title: "Association Analysis HW"
author: "Dean D'souza"
date: "July 7, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solutions:  

* First, install the package arules  

```{r,warning=FALSE,message=FALSE}
# Intalling the 'arules' package (commented out as already installed)
# install.packages("arules")
```

* Second, load the data set using the following command  

```{r,warning=FALSE,message=FALSE}
library(arules)
grd <- read.transactions("http://fimi.ua.ac.be/data/retail.dat", format="basket")
```

* Third, run the following commands and interpret the results  

```{r,warning=FALSE,message=FALSE}
itemFrequencyPlot(grd,support=.1)  #run with support .2, .3, & .5
summary(grd)
inspect(grd)  #you will have to stop the listing manually
```

* Create the rules object using apriori  

```{r,warning=FALSE,message=FALSE}
grdar <- apriori(grd,parameter=list(supp=.05,conf=.5))
inspect(grdar)
```

* Now that you have rules,.  
* Find a few interesting rules  
* Tell me something you learned from interpreting the rules  
* Show all your steps (especially in data conversion) using knitR  

```{r,warning=FALSE,message=FALSE}
```

* Next, tell what you would like to do next with the retail data  
  + Is there a hypothesis you would like to test?  
  + Is there data from another source you would like to add?  
  + Is there a predictive model you would like to build?  
  + Anything else?  
  + Tell me what would be interesting to you to do next  

```{r,warning=FALSE,message=FALSE}
```

* Tell me about project you would like to do with Association Analysis  
  + It can be a project at work  
  + Or, suppose you could download data from data.gov on healthcare, or education, or whatever  
  + What would you like to do Assocation Analysis on if you could  

```{r,warning=FALSE,message=FALSE}
```