---
title: 'Forecasting: Holt-Winters'
author: "Dean D'souza"
date: "August 2, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Setting up Nile Data
data("Nile")
head(Nile)

# Setting up the AirPassengers data
data("AirPassengers")
head(AirPassengers)

#Setting up the EUStockMarkets data
data("EuStockMarkets")
head(EuStockMarkets)

set.seed(695949)
```

## Solutions:

* Repeat the ts(), HoltWinters(), predict() and plot() functions on the Nile data as in these slides  
  
  We execute the R commands found in the lecture slides as follow:  
  
```{r,warning=FALSE,message=FALSE, fig.align='center',fig.height=4,fig.width=7}
# Setting up the nile data to be usable
nile.ts<-ts(Nile,start=1,frequency=10)
summary(nile.ts)

# Plotting the original data
plot(nile.ts)

#Creating the Holt-Winters object without gamma
nile.hw1<-HoltWinters(nile.ts,gamma = FALSE)
nile.hw1

# Plotting the Holt-Winters object
plot(nile.hw1)

# Predicting future values with the Holt-winters Object
nile.p1<-predict(nile.hw1, n.ahead=10)
ts.plot(nile.ts,nile.p1)

#Creating the new Holt-Winters object with gamma
nile.hw2<-HoltWinters(nile.ts)
nile.hw2

# Plotting the new Holt-Winters object
plot(nile.hw2)

# Predicting future values with the new Holt-winters Object
nile.p2<-predict(nile.hw2, n.ahead=10)
ts.plot(nile.ts,nile.p2)
```  

* Repeat the same functions with different values of alpha, beta, and gamma of your choosing on.  
  + AirPassengers  
  
  We perform the functions on the AirPassengers data set present in base R as follows:  
  
```{r,warning=FALSE,message=FALSE}
# Creating the Holt-Winters Object
ap.hw1<-HoltWinters(AirPassengers, alpha = 0.25,beta = FALSE, gamma = TRUE)
ap.hw1

# Plotting the Model
plot(ap.hw1)

# Predicting the Future values
ap.p1<-predict(ap.hw1, n.ahead=10)
ts.plot(AirPassengers,ap.p1)

# Creating the Holt-Winters Object such that the algorithm sets alpha, beta and gamma
ap.hw2<-HoltWinters(AirPassengers)
ap.hw2

# Plotting the Model
plot(ap.hw2)

# Predicting the Future values
ap.p2<-predict(ap.hw2, n.ahead=10)
ts.plot(AirPassengers,ap.p2)
```  
  
  In the first model created above we set alpha to 0.25 to get a closer approximation for predicting the values. We also set Beta to False to perform exponential smoothing. This is to done so as to maintain the upward trend seen in the data. We also set gamma to true in order to take into account the cyclical effect seen in the data. The second model gives us the actual alpha, beta and gamma values that should be used for a model with good fit.  
  
  + EuStockMarkets  
  + When using the EuStockMarkets, choose one column  
  
We perform the functions on the EUStockMarkets data set present in base R as follows (The column chosen for the purposes of this assignment is the UK FTSE column):

```{r,warning=FALSE,message=FALSE}
# Getting the data ready for Holt-Winters
FTSE<-EuStockMarkets[,4]
head(FTSE)

# Verifying the time series
plot(FTSE)

# Creating the Holt-Winters Object
FTSE.hw1<-HoltWinters(FTSE, alpha = 0.9 ,beta = FALSE , gamma = FALSE)
FTSE.hw1

# Plotting the Model
plot(FTSE.hw1)

# Predicting the Future values
FTSE.p1<-predict(FTSE.hw1, n.ahead=10)
ts.plot(FTSE,FTSE.p1)

# Creating the Holt-Winters Object such that the algorithm sets alpha, beta and gamma
FTSE.hw2<-HoltWinters(FTSE)
FTSE.hw2

# Plotting the Model
plot(FTSE.hw2)

# Predicting the Future values
FTSE.p2<-predict(FTSE.hw2, n.ahead=10)
ts.plot(FTSE,FTSE.p2)
```  
  
  In the first model created above we set alpha to 0.9 to get a closer approximation for predicting the values. We also set Beta to FALSE to perform exponential smoothing. We also set gamma to False as we do not want to fit a cyclical model (as the data does not show significant cyclical effects). The second model gives us the actual alpha, beta and gamma values that should be used for a model with good fit.  
  