---
title: "Cluster Analysis"
author: "Dean D'souza"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solutions:

* Using the mtcars data set

```{r,warning=FALSE,message=FALSE}
data(mtcars)
head(mtcars)
dat<-mtcars
remove(mtcars)
```  

* Create a kmeans object from the first, second, and third columns  

```{r,warning=FALSE,message=FALSE}

```  

* What is the size of each cluster?  
* What are the centers of each cluster?  
* What is the average disp, wt, and qsec of each cluster?  
* Describe each cluster in English.   
* Find a data set with at least 4 columns of numeric data and a categorical column  
* Run several scatter plots of the data  
* Create a kmeans object from the numeric data, you can pick K to be whatever you want  
* Determine the size of each cluster  
* Determine the centers of each cluster  
* Compare the clusters to the categorical data column as we did with the iris$Species column  
* For your chosen data set,   
  + Describe what each row of data represents  
  + Describe each of your columns used - give a one sentence description of the column  
  + If you know it, describe how the data was generated
* For the clusters  
  + Describe the size and means of clusters  
  + Give a one- or two-word description to each cluster - in other words, give each cluster a label or name  
* This is an exercise in turning your numeric data into something descriptive for non-statisticians  