---
title: "Cluster Analysis"
author: "Dean D'souza"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

## Solutions:

* __Using the mtcars data set__
  
  We setup the data as follows:  
  
```{r,warning=FALSE,message=FALSE}
data(mtcars)
head(mtcars)
dat<-mtcars
remove(mtcars)
set.seed(596979)
```  

* __Create a kmeans object from the first, second, and third columns__  
  
  We create the kmeans object after subsetting as follows:  

```{r,warning=FALSE,message=FALSE}
cars123<-mtcars[,1:3]
cars_km1<-kmeans(cars123, centers = 3)
```  

* __What is the size of each cluster?__  
  
  We can obtain the size of the three clusters as follows:  
  
```{r,message=FALSE,warning=FALSE}
cars_km1$size
```  

* __What are the centers of each cluster?__  
  
  We can obtain the centers of the clusters as follows:  

```{r,warning=FALSE,message=FALSE}
cars_km1$centers
```  

* __What is the average disp, wt, and qsec of each cluster?__  
  
  We obtain the details as follows:  

```{r,warning=FALSE,message=FALSE}
cars_478_avg<-aggregate(mtcars, by=list(cars_km1$cluster), FUN = mean)
cars_478_avg[,c(4,7,8)]
```

* __Describe each cluster in English.__   
  
  In the exercise above we have generated 3 clusters of sizes 14, 9 and 9 observations respectively. These clusters are centered around observations of seemingly increasing miles per gallon and decreasing displacement.

* __Find a data set with at least 4 columns of numeric data and a categorical column__  
  
  For the purpose of this assignment I chose to combine the first four columns of State.x77 and the State.region data into a single data frame as follows:
    
```{r,warning=FALSE,message=FALSE}
state<-na.omit(data.frame(state.x77[,1:4],state.region))
```  

* __Run several scatter plots of the data__  
    
    Please note in the following plots, the red line represents the regression line while the blue line represents the lowess line.  
    
```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.height=4,fig.width=6}
par(mfrow=c(2,2))

hist(state$Population)
hist(state$Income)
hist(state$Illiteracy)
hist(state$Life.Exp)
plot(state$state.region)

plot(state$Population,state$Income,main = "Population vs. Income",xlab = "Population",ylab = "Income")
abline(lm(state$Population~state$Income), col="red")
lines(lowess(state$Population,state$Income), col="blue")

plot(state$Population,state$Illiteracy,main = "Population vs. Illiteracy rate",xlab = "Population",ylab = "Illiteracy")
abline(lm(state$Population~state$Illiteracy), col="red")
lines(lowess(state$Population,state$Illiteracy), col="blue")

plot(state$Population,state$Life.Exp,main = "Population vs. Life Expectancy",xlab = "Population",ylab = "Life Expectancy")
abline(lm(state$Population~state$Life.Exp), col="red")
lines(lowess(state$Population,state$Life.Exp), col="blue")

plot(state$Income,state$Illiteracy,main = "Income vs. Illiteracy Rate",xlab = "Income",ylab = "Illiteract Rate")
abline(lm(state$Income~state$Illiteracy), col="red")
lines(lowess(state$Income,state$Illiteracy), col="blue")

plot(state$Income,state$Life.Exp,main = "Income vs. Life Expectancy",xlab = "Income",ylab = "Life Expectancy")
abline(lm(state$Income~state$Life.Exp), col="red")
lines(lowess(state$Income,state$Life.Exp), col="blue")

plot(state$Illiteracy,state$Life.Exp,main = "Illiteracy Rate vs. Life Expectancy",xlab = "Illiteracy Rate",ylab = "Life Expectancy")
abline(lm(state$Illiteracy~state$Life.Exp), col="red")
lines(lowess(state$Illiteracy,state$Life.Exp), col="blue")
```  

* __Create a kmeans object from the numeric data, you can pick K to be whatever you want__  
  
  We create the Kmeans object as follows:  

```{r,warning=FALSE,message=FALSE}
state1234<-state[,1:4]
state_km1<-kmeans(state1234,centers = 3)
```  

* __Determine the size of each cluster__  
  
  The size of the clusters are as follows:  

```{r,warning=FALSE,message=FALSE}
state_km1$size
``` 

* __Determine the centers of each cluster__  

  The centers of the clusters are as follows:  

```{r,warning=FALSE,message=FALSE}
state_km1$centers
```  

* __Compare the clusters to the categorical data column as we did with the iris$Species column__  
  
  We compare the clusters to the categorical data as follows:  
  
```{r,warning=FALSE,message=FALSE}
state_play<-state
state_play$cluster_id<-state_km1$cluster
head(state_play)
table(state_play$state.region,state_play$cluster_id)
```

* __For your chosen data set,__   
  + __Describe what each row of data represents__  
  
  Each row of the data set contains observations about the various states in the U.S.A. which includes average life expectancy, population, average income, percent illiteracy and region.  
  
  + __Describe each of your columns used - give a one sentence description of the column__  
    
  We now continue to describe the data with it's variables:
  
```{r,echo=FALSE,warning=FALSE,message=FALSE}
disc<-data.frame(c(1:5),c("Population","Income","Illiteracy","Life Exp.","state.region"),c("population estimate as of July 1, 1975","per capita income (1974)","illiteracy (1970, percent of population)","life expectancy in years (1969-71)","factor giving the region (Northeast, South, North Central, West) that each state belongs to."))
kable(disc,col.names = c("","Variable","Description"))
```
  
  + __If you know it, describe how the data was generated__  
    
    The State.x77 data set and state.region vector were combined together for the purpose of this assignment. These are provided with base R package of datasets. As all observations are arranged according to alphabetical order of states we only needed to combine the two normally.  
    
* __For the clusters__  
  + __Describe the size and means of clusters__  
    
    The clusters generated from the data consists of one relatively large cluster of size 27 observations, a medium cluster of 17 observations and a small cluster of 16 observations. We can also see from the table that a number of observations in the first cluster come from the Western states and a few from the other regions. Most of the southern states fall under cluster 3 whilw north central sates seem to favour clusters 1 and 3. Northeastern states seem to favour cluster 1.  
    Overall we can see that for the most part the clusters are formed over population densities and hence we name them in the following question appropriately.  
    
  + __Give a one- or two-word description to each cluster - in other words, give each cluster a label or name__  
    
    Cluster 1 <- Low Population  
    Cluster 2 <- High Population  
    Cluster 3 <- Medium Population    
